Usage metrics

### 2020-12-11 inst/analysis/usage_metrics.Rmd
usage metrics

The following is a set up chunk. Currently there is an variable "jan" 
used to run the script from docker on my machine. This should be
removed. (TODO)

@@

knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    if (before) { # record the current time before each chunk
      now <<- Sys.time()
    } else { # calculate the time difference after a chunk
      res <- difftime(Sys.time(), now)
      # return a character string to show the time
      paste("Time for this code chunk to run:", res, units(res))
}}}))
now <- Sys.time()
knitr::opts_chunk$set(
  echo = TRUE,
  fig.retina = 2,
  fig.width = 10,
  cache.lazy = FALSE,
  time_it = TRUE
)

@@

Read in the summarized calls data and the file that describes
which package belong to our corpus.

```{r read}
read_fst(eval_calls_path) %>% as_tibble() -> E_raw
read_fst(corpus_path) %>% as_tibble() -> C_raw
```

The two data sets are E and C for conciseness.  C is only used here to compute
the corpus size. E is used throughout. Recall that E_raw has the source data
with all columns.

```{r}
E_raw %>% select(
  run_package = package,        # The package which was run to trigger the eval
  ev_variant = eval_function,   # Which eval variant was called
  duplicates = nb_ev_calls,     # Number of identical eval calls (weight)
  src_ref = eval_call_srcref,   # Source ref for the eval call site
  file,                         # Script file name (name of this run)
  ev_package =eval_source,      # Package of the call site
  arg_size = expr_expression_length, # Size in bytes of eval argument string
  opcodes = direct_interpreter_eval  # Non-recursive AST nodes traversed
) -> E                               # by the interpreter

@@

Compute the number of times eval and it variants were called in the entire 
corpus. This is a dynamic count that aggregates all call sites within
the corpus.

```{r eval-calls-overview}
nb_calls <- (E %>% count(wt = duplicates))$n
eval_calls <- E %>%
  filter(ev_variant == "eval") %>%
  count(wt = duplicates)
evalq_calls <- E %>%
  filter(ev_variant == "evalq") %>%
  count(wt = duplicates)
eval.parent_calls <- E %>%
  filter(ev_variant == "eval.parent") %>%
  count(wt = duplicates)
local_calls <- E %>%
  filter(ev_variant == "local") %>%
  count(wt = duplicates)

r("nb eval and variant calls", nb_calls)
r("nb eval calls", eval_calls$n)
r("nb evalq calls", evalq_calls$n)
r("nb eval parent calls", eval.parent_calls$n)
r("nb local calls", local_calls$n)
```

There so mane calls: `r nb_calls`

# Number of eval call sites

```{r nb-eval-call-sites}
nb_ev_call_sites <- E %>%
  select(src_ref) %>%
  n_distinct(na.rm = TRUE)

call_sites_per_package <- E %>%
  group_by(ev_package) %>%
  summarize(nb_call_sites = n_distinct(src_ref, na.rm = TRUE))

nb_packages_one_call_site <- call_sites_per_package %>%
  filter(nb_call_sites == 1) %>%
  nrow()
packages_at_least_one_call_site <- call_sites_per_package %>%
  filter(nb_call_sites >= 1)
median_call_sites <- packages_at_least_one_call_site %>%
  summarise(med = median(nb_call_sites, na.rm = TRUE))
max_packages <- packages_at_least_one_call_site %>%
  filter(nb_call_sites == max(nb_call_sites, na.rm = TRUE)) %>%
  select(ev_package, nb_call_sites)
nb_no_calls_to_eval <- corpus_size -
  n_distinct(select(packages_at_least_one_call_site, ev_package))

# sites triggered over the whole corpus
r("nb eval call sites", nb_ev_call_sites) 
# How many packages in the corpus have a single triggered call site
r("nb one call site", nb_packages_one_call_site)
r("at least one call site", nrow(packages_at_least_one_call_site))
r("packages no calls to eval", nb_no_calls_to_eval)
r("nb median call sites", median_call_sites$med)
r("max call sites", max_packages$nb_call_sites)
r("max call sites package", max_packages$ev_package)
```

```{r call-sites-violin-plot}
E %>%
  group_by(ev_package) %>%
  mutate(nb_call_sites = n_distinct(src_ref, na.rm = TRUE)) %>%
  ggplot(aes(x = ev_variant, y = nb_call_sites, fill = ev_variant), na.rm = TRUE) +
  geom_violin(trim = FALSE, width = 1.7) +
  # geom_jitter(shape=16, position=position_jitter(0.2)) +
  labs(y = "Call sites per package") +
  theme(
    legend.position = c(0.5, 0.85),
    legend.box.background = element_rect(fill = "white", size = 0.1),
    legend.box = "horizontal",
    axis.title.x = element_blank()
  )

@@

<!-- How is the normalized distribution of eval calls?  -->

<!-- ```{r distribution-call-sites} -->
<!-- eval_calls_per_package %>% -->
<!--   ggplot(aes(x = n), na.rm = TRUE) + -->
<!--   geom_density(adjust = 10) + -->
<!--   scale_y_log10()  -->
<!-- ``` -->

<!-- (Like the violin plot, but smoothed) -->
(th)

>>>>

The setup chunk has a variable `jan` with setting to run the script on
docker on Jan's mamchine. This should be removed.

@@

knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    if (before) { # record the current time before each chunk
      now <<- Sys.time()
    } else { # calculate the time difference after a chunk
      res <- difftime(Sys.time(), now)
      # return a character string to show the time
      paste("Time for this code chunk to run:", round(res,2), units(res))
}}}))
now <- Sys.time()
knitr::opts_chunk$set(
  echo = TRUE,
  fig.retina = 2,
  fig.width = 10,
  cache.lazy = FALSE,
  time_it = TRUE
)

@@

Read in the summarized calls data (`E_raw`) and the file that describes
which package belong to our corpus (`C_raw`). For conciseness, `E` is
the input data set with only the columns that we use.

```{r read,echo=F,cache=F}
read_fst(eval_calls_path) %>% as_tibble() -> E_raw
read_fst(corpus_path) %>% as_tibble() -> C_raw
#read_csv(paste0(base_dir, "/run/package-evals-traced.9/package-eval-static.csv")) -> P

E_raw %>% select(
  run_package = package,        # The package which was run to trigger the eval
  ev_variant = eval_function,   # Which eval variant was called
  duplicates = nb_ev_calls,     # Number of identical eval calls (weight)
  src_ref = eval_call_srcref,   # Source ref for the eval call site
  file,                         # Script file name (name of this run)
  ev_package =eval_source,      # Package of the call site
  arg_size = expr_expression_length, # Size in bytes of eval argument string
  opcodes = direct_interpreter_eval  # Non-recursive AST nodes traversed
) -> E                               # by the interpreter

@@

# Eval Calls Overiew
The corpus has `r corpus_size` packages, and the data set has 
`r nrow(E)` distinct calls to the `eval` and its variants.

```{r eval-calls-overview,echo=F}
nb_calls <- E %>% 
  count(wt = duplicates) %>%
  .$n
eval_calls <- E %>%
  filter(ev_variant == "eval") %>%
  count(wt = duplicates)  %>%
  .$n
evalq_calls <- E %>%
  filter(ev_variant == "evalq") %>%
  count(wt = duplicates) %>%
  .$n
eval.parent_calls <- E %>%
  filter(ev_variant == "eval.parent") %>%
  count(wt = duplicates)  %>%
  .$n
local_calls <- E %>%
  filter(ev_variant == "local") %>%
  count(wt = duplicates) %>%
  .$n

TBL <- overview_table(
r("nb eval and variant calls", nb_calls),
r("nb eval calls", eval_calls),
r("nb evalq calls", evalq_calls),
r("nb eval parent calls", eval.parent_calls),
r("nb local calls", local_calls)
)
```

We recorded `r nb_calls` to eval, out of these, there were `r evalq_calls`
(`r fmt(ratio(evalq_calls,nb_calls))`) calls to `evalq`, `r eval.parent_calls` 
(`r fmt(ratio(evalq_calls,nb_calls))`) calls to `eval.parent`,  and
`r local_calls` (`r fmt(ratio(evalq_calls,nb_calls))`) calls to `local`.

`r TBL`

# Nb eval call sites

```{r nb-eval-call-sites, echo=F}
stopifnot(sum(is.na(E$src_ref))==0)  # No NAs in scr_ref!

sites <- E %>%             # src_ref indicates where the call originates from.
  pull(src_ref) %>% 
  n_distinct()             # This the number of distinct sites

sites_per_package <- E %>%
  count(ev_package, src_ref) %>%
  count(ev_package)        # All packages have at least one site!

max_packages <- sites_per_package %>%
  filter(n == max(n)) %>%
  .$n

TBL <- overview_table(
# Sites triggered over the whole corpus:
r("nb eval call sites", sites) ,
# How many packages in the corpus have a single triggered call site:
r("nb one call site", sites_per_package %>% filter(n==1)),  
r("at least one call site", sites_per_package),
r("packages no calls to eval", corpus_size - nrow(sites_per_package)),
r("nb median call sites",  median(sites_per_package$n)),
r("max call sites", max_packages),
)
```

`r TBL`

```{r call-sites-violin-plot, echo=F}
E %>%
  group_by(ev_package) %>%
  mutate(sites=n_distinct(src_ref)) %>%
  ggplot(aes(x=ev_variant, y=sites, fill=ev_variant), na.rm=TRUE) +
  geom_violin(trim=FALSE, width=1.7) +
  # geom_jitter(shape=16, position=position_jitter(0.2)) +
  labs(y="Call sites per package") +
  theme(
    legend.position=c(0.5, 0.85),
    legend.box.background=element_rect(fill="white", size=0.1),
    legend.box="horizontal",
    axis.title.x=element_blank()
  )

@@
